---
title: "Hemoglobin_project"
author: "Group K"
date: "October 4, 2018"
output: html_document
---

### Cohort selection for the Hemoglobin project 

```{r setup, include=FALSE}

library(bigrquery)
library(dplyr)
#library(plyr)
library(tidyr)
project <- "hst-953-2018"
library(knitr)
library (ggplot2)
library(Hmisc)
library(gridExtra)

```
#COHORT SELECTION

As MIMIC-III is primarily an intensive care unit (ICU) database, we will focus on patients admitted to and discharged from the ICU. We will select our cohort based off the ICU stay (identified by `icustay_id`). 

```{r}
#selcting the ICU admissions that have 

sql<- paste("
SELECT icu.subject_id, icu.hadm_id, icu.icustay_id, pat.gender,  
DATETIME_DIFF(adm.admittime, pat.dob, YEAR) AS ageatadmin,
RANK() OVER (PARTITION BY icu.subject_id ORDER BY icu.intime) AS icustay_id_order, 
DATETIME_DIFF(icu.outtime, icu.intime, HOUR) AS icu_length_of_stay, adm.deathtime, adm.admission_type, 
adm.admission_location, adm.ethnicity, adm.diagnosis, serv.prev_service, serv.curr_service
FROM `physionet-data.mimiciii_clinical.icustays` icu
INNER JOIN `physionet-data.mimiciii_clinical.admissions` adm
ON icu.hadm_id = adm.hadm_id
INNER JOIN `physionet-data.mimiciii_clinical.patients`  pat
ON icu.subject_id = pat.subject_id
INNER JOIN `physionet-data.mimiciii_clinical.services`  serv
ON icu.hadm_id = serv.hadm_id
")

Admissions <- query_exec(sql,project=project,use_legacy_sql=FALSE)


#patient exclusion critera
#Exluding patients younger than 15 or older tan 150 years
Admissions$age_exclusion<- ifelse(Admissions$age<15 | Admissions$age>150, 1, 0)
#Exluding patients second or extra admission, we are focusing in intial admisions
Admissions$stay_exclusion<- ifelse(Admissions$icustay_id_order >1, 1, 0)
#Exluding patients whit less than 4 hours of ICU stay 
Admissions$lenght_stay_exclusion<- ifelse(Admissions$icu_length_of_stay <4, 1, 0)

#total Exclusion
Admissions$exclude <- ifelse(Admissions$age_exclusion | Admissions$stay_exclusion | Admissions$lenght_stay_exclusion, 1, 0)

Admission_final <- subset(Admissions, Admissions$exclude ==0)

table(Admissions$age_exclusion)
table(Admissions$stay_exclusion)
table(Admissions$lenght_stay_exclusion)



#detach(package:plyr)
#since the admission can have multiple services on charge I concatenate them 
Admission_final <- Admission_final %>%
                   group_by(icustay_id) %>%
                   mutate(current_service = paste(curr_service, collapse=", ")) %>%
                   select("subject_id","hadm_id", "icustay_id", "gender", "ageatadmin",
                                  "icu_length_of_stay", "admission_type",  "admission_location", 
                                  "ethnicity", "diagnosis","current_service") %>% 
                    unique()

```
There is a total of `nrow(Admission_final)` valid admission

##SELECTING THE HEMOGLOBINE VALUES AT BASELINE AND AT INITIAL TIMEPOINTS 
```{r}

#### evaluating Laboratories ---

sql <- paste("SELECT * FROM `physionet-data.mimiciii_clinical.d_labitems`")
labsnametable<- query_exec(sql,project=project,use_legacy_sql=FALSE)

labsnametable[grep('Hemo',labsnametable$LABEL),]

labsnametable[grep('B12',labsnametable$LABEL),]
labsnametable[grep('50810',labsnametable$ITEMID),]

```

ITEMIDs 50811, 50855 and 51222 are hemoglobin codes - selected by LOIC_CODE 718-7
###SELECTING PATIENT THAT HAVE A BASELINE VALUE AND ARE NOT EXCLUDED BY OUR PROVIUS SELECTION CRITERIA. 
```{r}

sql <- paste("
WITH hb AS
(SELECT labs.row_id, labs.subject_id,icu.icustay_id, labs.hadm_id, labs.charttime, adm.admittime, 
            labs.itemid, labs.valuenum, namelab.label,
            DATETIME_DIFF(labs.charttime, adm.admittime, DAY) AS labs_timelag
            FROM `physionet-data.mimiciii_clinical.labevents` labs
             INNER JOIN `physionet-data.mimiciii_clinical.d_labitems` namelab
              ON labs.itemid = namelab.itemid
            INNER JOIN `physionet-data.mimiciii_clinical.admissions` adm
            ON labs.subject_id = adm.subject_id
            INNER JOIN `physionet-data.mimiciii_clinical.icustays` icu
            ON adm.hadm_id = icu.hadm_id
            WHERE labs.itemid in (50811, 50855, 51222))
SELECT hb.subject_id, hb.icustay_id, hb.hadm_id, hb.charttime, hb.admittime, hb.itemid, 
            hb.valuenum, hb.label, hb.labs_timelag FROM hb 
            WHERE (hb.labs_timelag > -360 AND  hb.labs_timelag < -3)
  ")


Hemo_baseline_all <- query_exec(sql,project=project,use_legacy_sql=FALSE,max_pages = Inf)

#selecting patients that match the ICUstay ID
Hemo_baseline<- Hemo_baseline_all %>% filter(icustay_id %in% Admission_final$icustay_id ) 

#Removing the values of hemoglobin that are NA = total of 546
Hemo_baseline<-subset(Hemo_baseline, ! is.na(Hemo_baseline$valuenum))

 
#Selecting the most recent Hb value
Hemo_baseline<- Hemo_baseline %>% 
                group_by(icustay_id) %>%
                slice(which.max(labs_timelag))%>%
                select(subject_id, icustay_id, charttime, admittime, valuenum, labs_timelag) %>%
                rename(Hb_baseline_time = charttime, Hb_value_baseline = valuenum,
                       Hem_base_timelag = labs_timelag)

summary(Hemo_baseline$Hem_base_timelag)
hist(Hemo_baseline$Hem_base_timelag)

```

THERE ARE `length(unique(Hemo_baseline$subject_id))` PATIENTS WITH AT LEAST ONE VALUE BEFORE ADMISSION 

#SELECTING THE INITIAL HEMOGLOBIN VALUE FOR THE COHORT OF PATIENTS 
```{r}
#looking for outlayers before selecting the right summary value

boxplot(Hemo_baseline$Hb_value_baseline, lab="Hemoglobin baseline Result")


### table of the patients that have HEMOGLOBIN values below 5 to asses if they are really outlayer or not. 
for (i in 1:length(unique(subset(Hemo_baseline, Hb_value_baseline<5)$subject_id))){
  table<-Hemo_baseline_all %>% 
  filter(subject_id==unique(subset(Hemo_baseline, Hb_value_baseline<5)$subject_id)[i]) %>%
  arrange(labs_timelag) %>%
  select(subject_id,icustay_id,label,valuenum,labs_timelag) %>%
  knitr::kable(caption=paste("Extreme Outlier hemoglobin, ID:", subset(Hemo_baseline, Hb_value_baseline<5)$subject_id[i]))
  print(table)
}


### table of the patients that have HEMOGLOBIN values above 18 to asses if they are really outlayer or not. 
for (i in 1:length(unique(subset(Hemo_baseline, Hb_value_baseline>18)$subject_id))){
  table<-Hemo_baseline_all %>% 
  filter(subject_id==unique(subset(Hemo_baseline, Hb_value_baseline>18)$subject_id)[i]) %>%
  arrange(labs_timelag) %>%
  select(subject_id,icustay_id,label,valuenum,labs_timelag) %>%
  knitr::kable(caption=paste("Extreme Outlier hemoglobin, ID:", subset(Hemo_baseline, Hb_value_baseline>18)$subject_id[i]))
  print(table)
}



```
There are values of hemoglobing below 5mg/dl that might be considered otlyers and and values over 18mg/dl that can be consider outsie too. there are not too many and looking at the previous values they seem reasonable. No patient was removed because of outlayers. 

 

```{r}
sql <- paste("
WITH hb AS
(SELECT labs.row_id, labs.subject_id,icu.icustay_id, icu.intime, labs.hadm_id, labs.charttime, 
            adm.admittime, labs.itemid, labs.valuenum, namelab.label,
            DATETIME_DIFF(labs.charttime ,adm.admittime, MINUTE) AS labs_minutelag,
            DATETIME_DIFF(labs.charttime ,icu.intime, MINUTE) AS labs_icuminutelag
            FROM `physionet-data.mimiciii_clinical.labevents` labs
             INNER JOIN `physionet-data.mimiciii_clinical.d_labitems` namelab
              ON labs.itemid = namelab.itemid
            INNER JOIN `physionet-data.mimiciii_clinical.admissions` adm
            ON labs.subject_id = adm.subject_id
            INNER JOIN `physionet-data.mimiciii_clinical.icustays` icu
            ON adm.hadm_id = icu.hadm_id
            WHERE labs.itemid in (50811, 50855, 51222))
SELECT hb.subject_id, hb.icustay_id, hb.hadm_id, hb.charttime, hb.admittime, hb.itemid, 
            hb.valuenum, hb.label, hb.labs_icuminutelag, hb.labs_minutelag FROM hb
  WHERE (hb.labs_minutelag >= -1440 AND hb.labs_icuminutelag <= 1440 )
  ")


Hemo_initial_all <- query_exec(sql,project=project,use_legacy_sql=FALSE,max_pages = Inf)

#selecting patients that match the ICUstay ID
Hemo_initial<- Hemo_initial_all %>% filter(icustay_id %in% Admission_final$icustay_id ) 

#Removing the values of hemoglobin that are NA = total of 903
sum(is.na(Hemo_initial$valuenum))
Hemo_initial<-subset(Hemo_initial_all, ! is.na(Hemo_initial_all$valuenum))
#Removing the values of hemoglobin that are 0 
Hemo_initial<-subset(Hemo_initial, ! Hemo_initial$valuenum==0)


#Selecting the Hb value closest to the admission time 
Hemo_initial<- Hemo_initial %>% 
                    group_by(icustay_id) %>%
                     slice(which.min(abs(labs_minutelag)))

#distribution of times when the Hb was taken and reported 
hist(Hemo_initial$labs_minutelag)

#selecting subgroups of data based on ICU in admission time 
Hemo_initial24h<- Hemo_initial %>%
                      select(subject_id, icustay_id, charttime, valuenum) %>%
                      rename(Hb_initial_time = charttime, Hb_value_initial = valuenum)

Hemo_initial6h<- Hemo_initial %>%
                      filter (labs_icuminutelag <=360) %>%
                      select(subject_id, icustay_id, charttime, valuenum) %>%
                      rename(Hb_initial_time = charttime, Hb_value_initial = valuenum)

Hemo_initial4h<- Hemo_initial %>%
                      filter (labs_icuminutelag <=240) %>%
                      select(subject_id, icustay_id, charttime, valuenum) %>%
                      rename(Hb_initial_time = charttime, Hb_value_initial = valuenum)

Hemo_initial2h<- Hemo_initial %>%
                      filter (labs_icuminutelag <=120) %>%
                      select(subject_id, icustay_id, charttime, valuenum) %>%
                      rename(Hb_initial_time = charttime, Hb_value_initial = valuenum)


#looking for outlayers before selecting the right summary value

boxplot(Hemo_initial24h$Hb_value_initial, lab="Hemoglobin initial Result")

```

### DATABASE WITH THE BASELINE AND INITAL HEMOGLOBIN VALUES FOR THE PATIENTS THAT HAVE OUR REQUIRED SELECTION CRITERIA  
```{r}

### merging the data set 

Hb_cohort24h<- merge(Hemo_baseline, Hemo_initial24h, by=c("subject_id", "icustay_id"))
Hb_cohort6h<- merge(Hemo_baseline, Hemo_initial6h, by=c("subject_id", "icustay_id"))
Hb_cohort4h<- merge(Hemo_baseline, Hemo_initial4h, by=c("subject_id", "icustay_id"))
Hb_cohort2h<- merge(Hemo_baseline, Hemo_initial2h, by=c("subject_id", "icustay_id"))

boxplot(Hb_cohort24h$Hb_value_baseline, Hb_cohort24h$Hb_value_initial, lab="Hemoglobin Result")

```

### ADDING THE INFORMATION ABOUT DEMOGRAOHICS 

```{r}
Masterfile24h<- merge(Hb_cohort24h, Admission_final, by=c("subject_id", "icustay_id"), all.x = TRUE)
Masterfile6h<- merge(Hb_cohort6h, Admission_final, by=c("subject_id", "icustay_id"), all.x = TRUE)
Masterfile4h<- merge(Hb_cohort4h, Admission_final, by=c("subject_id", "icustay_id"), all.x = TRUE)
Masterfile2h<- merge(Hb_cohort2h, Admission_final, by=c("subject_id", "icustay_id"), all.x = TRUE)

```

### SELECTING OTHER RELEVANT LABS 
```{r}
#labs of interest based on the google doc document 
labs_of_interest<- "(51221, 50810, 51300, 51301, 51265, 50862, 50885, 50882, 50803, 50893, 50902, 50912,
                      50924,50925, 51237, 50952, 50953, 50813, 50954, 51250, 51248, 51249, 50818, 50820,
                      50970, 50821, 50971, 50976, 51274, 51275, 51279, 51006, 50983, 50998, 51010)"
sql <- paste("
WITH all_labs AS
(SELECT labs.subject_id,icu.icustay_id, icu.intime, labs.hadm_id, labs.charttime, 
            adm.admittime, labs.itemid, labs.valuenum, namelab.label,
            DATETIME_DIFF(labs.charttime ,adm.admittime, MINUTE) AS labs_minutelag, 
            DATETIME_DIFF(labs.charttime ,icu.intime, MINUTE) AS labs_icuminutelag
            FROM `physionet-data.mimiciii_clinical.labevents` labs
             INNER JOIN `physionet-data.mimiciii_clinical.d_labitems` namelab
              ON labs.itemid = namelab.itemid
            INNER JOIN `physionet-data.mimiciii_clinical.admissions` adm
            ON labs.subject_id = adm.subject_id
            INNER JOIN `physionet-data.mimiciii_clinical.icustays` icu
            ON adm.hadm_id = icu.hadm_id
            WHERE labs.itemid in", labs_of_interest, ")
SELECT all_labs.subject_id, all_labs.icustay_id, all_labs.hadm_id, all_labs.charttime, all_labs.admittime,
            all_labs.itemid, all_labs.valuenum, all_labs.label, all_labs.labs_minutelag, all_labs.labs_icuminutelag
            FROM all_labs
            WHERE (all_labs.labs_minutelag >= -1440 AND  all_labs.labs_icuminutelag <= 1440)
  ")


All_labs <- query_exec(sql,project=project,use_legacy_sql=FALSE,max_pages = Inf)

###########################IM HERE#######################################
#selecting patients that match the ICUstay ID
All_labs<- All_labs %>% filter(icustay_id %in% Admission_final$icustay_id)

#recoding labs name that have multiple ITEMID codes
All_labs$label <- recode (All_labs$label, "Calculated Bicarbonate, Whole Blood"= "Bicarbonate", 
                          "WBC Count" = "White Blood Cells", "Hematocrit, Calculated"= "Hematocrit")

#Removing values that are 0
All_labs$valuenum[All_labs$valuenum==0] <- NA
#Removing the values that are NA
sum(is.na(All_labs$valuenum))
All_labs<-subset(All_labs, ! is.na(All_labs$valuenum))

#Selecting the most recent lab value
All_labs24h<- All_labs %>% 
                group_by(icustay_id, label) %>%
                slice(which.min(abs(labs_minutelag))) %>%
                select(subject_id, icustay_id, label, valuenum)%>%
                spread(key=label, value=valuenum)

summary(All_labs2)


#Please note in the summary of the data that some of the max and min values are unrealistics and we will have to make adjustments for posible outliers. 

##chechink for outliers
for (i in 1:length(unique(subset(All_labs24h, Albumin>5.5)$subject_id))) {
  table<-All_labs %>% 
  filter(label == "Albumin", subject_id==unique(subset(All_labs24h, Albumin>5.5)$subject_id)[i]) %>%
  arrange(labs_minutelag) %>%
  select(subject_id,icustay_id,label,valuenum,labs_minutelag) %>%
  knitr::kable(caption=paste("Extreme Outlier Albumin, ID:", unique(subset(All_labs24h,
                                                                           Albumin>5.5)$subject_id[i])))
  print(table)
}

for (i in 1:length(unique(subset(All_labs24h, `Bilirubin, Total`>60)$subject_id))) {
  table<-All_labs %>% 
  filter(label == "Bilirubin, Total", subject_id==unique(subset(All_labs24h,`Bilirubin, Total`>60)$subject_id)[i]) %>%
  arrange(labs_minutelag) %>%
  select(subject_id,icustay_id,label,valuenum,labs_minutelag) %>%
  knitr::kable(caption=paste("Extreme Outlier Albumin, ID:", unique(subset(All_labs24h,
                                                                           `Bilirubin, Total`>60)$subject_id[i])))
  print(table)
}

All_labs24h$Ferritin

for (i in 1:length(unique(subset(All_labs24h, Ferritin>3000)$subject_id))) {
  table<-All_labs %>% 
  filter(label == "Ferritin", subject_id==unique(subset(All_labs24h,Ferritin>3000)$subject_id)[i]) %>%
  arrange(labs_minutelag) %>%
  select(subject_id,icustay_id,label,valuenum,labs_minutelag) %>%
  knitr::kable(caption=paste("Extreme Outlier Albumin, ID:", unique(subset(All_labs24h,
                                                                           Ferritin>3000)$subject_id[i])))
  print(table)
}



#selecting subgroups of data based on ICU in admission time 
Labs_6h<- All_labs %>%
          filter(labs_icuminutelag<=360) %>%
          group_by(icustay_id, label) %>%
          slice(which.min(abs(labs_minutelag))) %>%
          select(subject_id, icustay_id, label, valuenum)%>%
          spread(key=label, value=valuenum)


Labs_4h<- All_labs %>%
          filter(labs_icuminutelag<=240) %>%
          group_by(icustay_id, label) %>%
          slice(which.min(abs(labs_minutelag))) %>%
          select(subject_id, icustay_id, label, valuenum)%>%
          spread(key=label, value=valuenum)

Labs_2h<- All_labs %>%
          filter(labs_icuminutelag<=120) %>%
          group_by(icustay_id, label) %>%
          slice(which.min(abs(labs_minutelag))) %>%
          select(subject_id, icustay_id, label, valuenum)%>%
          spread(key=label, value=valuenum)


```
### ADDING THE INFORMATION ABOUT OTHER LABORATORY VALUES  

```{r}
Masterfile24h<- merge(Masterfile24h, All_labs24h, by=c("subject_id", "icustay_id"), all.x = TRUE)
Masterfile6h<- merge(Masterfile6h, Labs_6h, by=c("subject_id", "icustay_id"), all.x = TRUE)
Masterfile4h<- merge(Masterfile4h, Labs_4h, by=c("subject_id", "icustay_id"), all.x = TRUE)
Masterfile2h<- merge(Masterfile2h, Labs_2h, by=c("subject_id", "icustay_id"), all.x = TRUE)


```
### SELECTING vITAL SIGNS WITHING THE FIRTS DAY OF ICU ADMISSION 
```{r}
#Vitals of interest based on the google doc document 
vitals_of_interest<- "(211, 220045, 51, 442, 455, 6701, 220179, 220050, 8368, 8440, 8441, 8555, 220180,
                      220051, 456, 52, 6702, 443, 220052, 220181, 225312, 618, 615, 220210, 224690,
                      646, 220277, 807, 811, 1529, 3745, 3744, 225664, 220621, 226537, 223762, 676,
                      223761, 678) "
sql <- paste("
WITH vitals AS
(SELECT ce.subject_id,icu.icustay_id, icu.intime, ce.hadm_id, ce.charttime, 
            adm.admittime, ce.itemid, ce.valuenum, namece.label,
            DATETIME_DIFF(ce.charttime, icu.intime, MINUTE) AS ce_minutelag
            FROM `physionet-data.mimiciii_clinical.chartevents` ce
             INNER JOIN `physionet-data.mimiciii_clinical.d_items` namece
              ON ce.itemid = namece.itemid
            INNER JOIN `physionet-data.mimiciii_clinical.admissions` adm
            ON ce.subject_id = adm.subject_id
            INNER JOIN `physionet-data.mimiciii_clinical.icustays` icu
            ON adm.hadm_id = icu.hadm_id
            WHERE ce.itemid in", vitals_of_interest ,")
SELECT vitals.subject_id, vitals.icustay_id, vitals.hadm_id, vitals.charttime, vitals.admittime,
            vitals.itemid, vitals.valuenum, vitals.label, vitals.ce_minutelag FROM vitals
  WHERE (vitals.ce_minutelag >= -1440 AND  vitals.ce_minutelag <= 1440)
  ")


vitals <- query_exec(sql,project=project,use_legacy_sql=FALSE,max_pages = Inf)

#selecting patients that match the ICUstay ID
vitals<- vitals %>% filter(icustay_id %in% Admission_final$icustay_id)

#Recoding Temperature values to centigrades
vitals$values <- ifelse(vitals$itemid %in% c(223761,678), (vitals$valuenum-32)/1.8, vitals$valuenum)


#recoding vitals name that have multiple ITEMID codes
vitals$LABELS <- ifelse(vitals$itemid %in% c(211,220045), "HR", 
                        ifelse(vitals$itemid %in% c(51,442,455,6701,220179,220050), "SBP",
                        ifelse(vitals$itemid %in% c(8368,8440,8441,8555,220180,220051), "DBP",
                        ifelse(vitals$itemid %in% c(456,52,6702,443,220052,220181,225312),"MAP", 
                        ifelse(vitals$itemid %in% c(618,615,220210,224690), "RR", 
                        ifelse(vitals$itemid %in% c(646,220277), "SPO2", 
                        ifelse(vitals$itemid %in% c(807,811,1529,3745,3744,225664,220621,226537), "GLU",
                        ifelse(vitals$itemid %in% c(223762,676,223761,678), "TEMP", "ERROR"))))))))
                        
  
#Removing EXTREM VALUEES - the values that are NA = total 1606
vitals$values <- ifelse(vitals$LABELS=="HR" & (vitals$valuenum<=0|vitals$valuenum>300), NA, 
                        ifelse(vitals$LABELS=="SBP" & (vitals$valuenum<=0|vitals$valuenum>400), NA,
                        ifelse(vitals$LABELS=="DBP" & (vitals$valuenum<=0|vitals$valuenum>300), NA,
                        ifelse(vitals$LABELS=="MAP" & (vitals$valuenum<=0|vitals$valuenum>300), NA,
                        ifelse(vitals$LABELS=="RR" & (vitals$valuenum<=0|vitals$valuenum>70), NA,
                        ifelse(vitals$LABELS=="SPO2" & (vitals$valuenum<=0|vitals$valuenum > 100), NA,
                        ifelse(vitals$LABELS=="GLU" & (vitals$valuenum<=0|vitals$valuenum>1000), NA,
                        ifelse(vitals$LABELS=="TEMP" & (vitals$values<=0|vitals$values>50), NA,
                               vitals$values))))))))

sum(is.na(vitals$valuenum))
sum(is.na(vitals$values))

vitals<-subset(vitals, ! is.na(vitals$values))
  
#Summarizing the values of the vitals signs 

vitals24<- vitals %>%
          select(subject_id, icustay_id, LABELS, values) %>%
          group_by(icustay_id, LABELS) %>%
          summarise(mean=mean(values), sd=sd(values), max=max(values), min=min(values)) %>% 
          gather(variable, value, -(icustay_id:LABELS)) %>%
          unite(temp, LABELS, variable) %>%
          spread(temp, value)

vitals6<- vitals %>%
          filter(ce_minutelag<=360) %>%
          select(subject_id, icustay_id, LABELS, values) %>%
          group_by(icustay_id, LABELS) %>%
          summarise(mean=mean(values), sd=sd(values), max=max(values), min=min(values)) %>% 
          gather(variable, value, -(icustay_id:LABELS)) %>%
          unite(temp, LABELS, variable) %>%
          spread(temp, value)

vitals4<- vitals %>%
          filter(ce_minutelag<=240) %>%
          select(subject_id, icustay_id, LABELS, values) %>%
          group_by(icustay_id, LABELS) %>%
          summarise(mean=mean(values), sd=sd(values), max=max(values), min=min(values)) %>% 
          gather(variable, value, -(icustay_id:LABELS)) %>%
          unite(temp, LABELS, variable) %>%
          spread(temp, value)

vitals2<- vitals %>%
          filter(ce_minutelag<=120) %>%
          select(subject_id, icustay_id, LABELS, values) %>%
          group_by(icustay_id, LABELS) %>%
          summarise(mean=mean(values), sd=sd(values), max=max(values), min=min(values)) %>% 
          gather(variable, value, -(icustay_id:LABELS)) %>%
          unite(temp, LABELS, variable) %>%
          spread(temp, value)

    
summary(vitals2)
```


### ADDING THE INFORMATION ABOUT OTHER VITALS TO THE MASTER FILE   

```{r}
Masterfile24h<- merge(Masterfile24h, vitals24, by="icustay_id", all.x = TRUE)
Masterfile6h<- merge(Masterfile6h, vitals6, by="icustay_id", all.x = TRUE)
Masterfile4h<- merge(Masterfile4h, vitals4, by="icustay_id", all.x = TRUE)
Masterfile2h<- merge(Masterfile2h, vitals2, by="icustay_id", all.x = TRUE)


```


### SELECTING VASOPRESSORS 

```{r}
#Vitals of interest based on the google doc document 

vassopressors_mv<- "(221906, 221749, 222315, 221289, 221986, 221662, 221653) "

sql <- paste("
SELECT  icu.subject_id, icu.icustay_id, icu.intime, mv.itemid, mv.starttime, mv.endtime, mv. rate,
DATETIME_DIFF(mv.starttime, icu.intime, MINUTE) AS ce_minutelag
FROM `physionet-data.mimiciii_clinical.inputevents_mv` mv
INNER JOIN `physionet-data.mimiciii_clinical.icustays` icu
ON mv.icustay_id = icu.icustay_id
WHERE itemid in", vassopressors_mv)

vassopressors <- query_exec(sql,project=project,use_legacy_sql=FALSE,max_pages = Inf)

#selecting patients that match the ICUstay ID , and filtering time of event 

vassopressors<- vassopressors %>% filter(icustay_id %in% Admission_final$icustay_id, ce_minutelag<=1440)

vp24<- vassopressors %>%
          select(subject_id, icustay_id) %>%
          group_by(icustay_id) %>%
          mutate(vassopresor_doses= n()) 

vp6<- vassopressors %>%
          filter(ce_minutelag<=360) %>%
          select(subject_id, icustay_id) %>%
          group_by(icustay_id) %>%
          mutate(vassopresor_doses= n()) 

vp4<- vassopressors %>%
          filter(ce_minutelag<=240) %>%
          select(subject_id, icustay_id) %>%
          group_by(icustay_id) %>%
          mutate(vassopresor_doses= n()) 
vp2<- vassopressors %>%
          filter(ce_minutelag<=120) %>%
          select(subject_id, icustay_id) %>%
          group_by(icustay_id) %>%
          mutate(vassopresor_doses= n()) 


```

### ADDING THE INFORMATION ABOUT OTHER VITALS TO THE MASTER FILE   

```{r}
Masterfile24h<- merge(Masterfile24h, vp24, by=c("subject_id", "icustay_id"), all.x = TRUE)
Masterfile6h<- merge(Masterfile6h, vp6, by=c("subject_id", "icustay_id"), all.x = TRUE)
Masterfile4h<- merge(Masterfile4h, vp4, by=c("subject_id", "icustay_id"), all.x = TRUE)
Masterfile2h<- merge(Masterfile2h, vp2, by=c("subject_id", "icustay_id"), all.x = TRUE)


Masterfile24h$Bleed_dx <- ifelse(grepl("BLEED",Masterfile24h$diagnosis)| grepl("HEMORRHAGE",Masterfile24h$diagnosis),
                                 "yes", "No")

Masterfile6h$Bleed_dx <- ifelse(grepl("BLEED",Masterfile6h$diagnosis)| grepl("HEMORRHAGE",Masterfile6h$diagnosis),
                                 "yes", "No")

Masterfile4h$Bleed_dx <- ifelse(grepl("BLEED",Masterfile4h$diagnosis)| grepl("HEMORRHAGE",Masterfile4h$diagnosis),
                                 "yes", "No")

Masterfile2h$Bleed_dx <- ifelse(grepl("BLEED",Masterfile2h$diagnosis)| grepl("HEMORRHAGE",Masterfile2h$diagnosis),
                                 "yes", "No")


write.csv(Masterfile24h, "C:/Users/ch162944/Dropbox (Berde Lab)/HST.953/Masterfile24h.csv" )
write.csv(Masterfile6h, "C:/Users/ch162944/Dropbox (Berde Lab)/HST.953/Masterfile6h.csv" )
write.csv(Masterfile4h, "C:/Users/ch162944/Dropbox (Berde Lab)/HST.953/Masterfile4h.csv" )
write.csv(Masterfile2h, "C:/Users/ch162944/Dropbox (Berde Lab)/HST.953/Masterfile2h.csv" )




```
